### 2023-11-17 ms.qmd
Updates and graphs

pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, MASS)

@@

# Divide Treatment in CORT and Temperature
data <- data %>%
      mutate(temp = gsub("[AB]_", "", trt),
          cort = gsub("_[2][38]", "", trt))  %>% 
  data.frame()

# MASS
# Create new database to analyse increases in mass (i.e. having mass after and before in different columns)
data_2 <- pivot_wider(data, id_cols=c(lizard,temp,cort,species,food_ingested), names_from = Day, values_from = mass)
data_2 <- data_2 %>% mutate(delta_mass=After-Before) 
data_2 <- data_2 %>% mutate(rescaled_delta_mass = delta_mass - min(delta_mass) + 1)
write.csv(data_2, "./data/Stressor_2.csv")
# Transform mass into mg in new database
data_2$mass <- data$mass*1000

# Estimate mass increase

# Split data by species in new database
deli_stressor_2 <- data_2 %>% filter (species == "delicata")
guich_stressor_2 <- data_2 %>% filter (species == "guichenoti")

# Explore distribution of mass increase
  # A) L.deli
plot_mass <- ggplot(deli_stressor_2, aes(x = mass, fill = interaction(as.factor(temp),as.factor(cort), sep="_"))) +
  geom_histogram(position = "identity", alpha = 0.5, bins = 30, color = "black") +
  scale_fill_manual(values = c("28_A" = "red", "23_A" = "blue", "28_B" = "salmon", "23_B" = "lightblue")) +
  labs(x = "Mass", y = "Density", fill = "Temperature ~ CORT") +
  theme_minimal() +
  facet_grid(cort*temp~Day, scales = "free_y")
print(plot_mass)

mass_normal_test<-by(deli_stressor_2$mass, interaction (deli_stressor$cort, deli_stressor$temp, deli_stressor$Day), function(x) shapiro.test((x)))
print(mass_normal_test)

  #B) L.guich
plot_mass <- ggplot(guich_stressor_2, aes(x = mass, fill = interaction(as.factor(temp),as.factor(cort), sep="_"))) +
  geom_histogram(position = "identity", alpha = 0.5, bins = 30, color = "black") +
  scale_fill_manual(values = c("28_A" = "red", "23_A" = "blue", "28_B" = "salmon", "23_B" = "lightblue")) +
  labs(x = "Mass", y = "Density", fill = "Temperature ~ CORT") +
  theme_minimal() +
  facet_grid(cort*temp~Day, scales = "free_y")
print(plot_mass)

mass_normal_test<-by(guich_stressor$mass, interaction (guich_stressor$cort, guich_stressor$temp, guich_stressor$Day), function(x) shapiro.test((x)))
print(mass_normal_test)

@@

```{r, models1}
#| label: models1
#| echo: false
#| warning: false

# Generalized Mixed Models delicata
GMM_deli <- glmer(FC_associative ~ Associative_Trial*temp*cort + group*Associative_Trial + (1 + Associative_Trial|lizard_id), data = deli_associative, family = binomial(link = "logit"))
summary(GMM_deli)

emmeans(GMM_deli, pairwise ~ temp*Associative_Trial)
emmeans(GMM_deli, pairwise ~ cort*Associative_Trial)

# Generalized Mixed Models guichenoti
GMM_guich <- glmer(FC_associative ~ Associative_Trial*temp*cort + group*Associative_Trial + (1 + Associative_Trial|lizard_id), data = guich_associative, family = binomial(link = "logit"))
summary(GMM_guich)

emmeans(GMM_guich, pairwise ~ temp*Associative_Trial)
emmeans(GMM_guich, pairwise ~ cort*Associative_Trial)
emmeans(GMM_guich, pairwise ~ temp*cort*Associative_Trial)

```{r, models2}
#| label: models2
#| echo: false
#| warning: false

# Bayesian model delicata
  deli_mod <- brm(FC_associative ~ cort*temp*Associative_Trial + group*Associative_Trial + (1 + Associative_Trial|lizard_id), data = deli_associative, family = bernoulli(link = "logit"), chains = 4, cores = 4, iter = 2000, warmup = 1000, control = list(adapt_delta = 0.99))
  summary(deli_mod)
  plot(deli_mod)

# Bayesian model guichenoti
  guich_mod <- brm(FC_associative ~ cort*temp*Associative_Trial + group*Associative_Trial + (1 + Associative_Trial|lizard_id), data = guich_associative, family = bernoulli(link = "logit"), chains = 4, cores = 4, iter = 2000, warmup = 1000, control = list(adapt_delta = 0.99))
  summary(deli_mod)
# R2
bayes_R2(mod)

@@

```{r, plots}
#| label: plots
#| echo: false
#| warning: false

# Plot delicata
plot_deli <- ggplot(deli_associative, aes(x = Associative_Trial, y = FC_associative, color=interaction(cort,temp))) +  
  geom_smooth(method = "glm", formula = y ~ x, method.args = list(family = "binomial"),na.action = na.exclude) +
  facet_grid(as.factor(deli_associative$group)) + 
  labs(y = "Colour Association Task", x = "Trial")  
print(plot_deli)

# Plot guichenoti
plot_guich <- ggplot(guich_associative, aes(x = Associative_Trial, y = FC_associative, color=interaction(cort,temp))) +  
  geom_smooth(method = "glm", formula = y ~ x, method.args = list(family = "binomial"),na.action = na.exclude) +
  facet_grid(as.factor(guich_associative$group)) + 
  labs(y = "Colour Association Task", x = "Trial") 
print(plot_guich)
```

>>>>

pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, MASS, dplyr, broom)

@@

# Divide Treatment in CORT and Temperature Transform mass into mg
data <- data %>%
      mutate(temp = gsub("[AB]_", "", trt),
          cort = gsub("_[2][38]", "", trt))  %>% 
      mutate(mass=mass*1000) %>%
data.frame()
data$cort <- factor(data$cort, 
        levels = c("A" ,"B"),
        labels = c("Control", "CORT"))
data$temp <- factor(data$temp,
        levels = c("23", "28"),
        labels = c("Cold", "Hot"))
data$Day <- factor(data$Day,
        levels = c("Before", "After"))

# MASS
# Create new database to analyse increases in mass (i.e. having mass after and before in different columns).
data_2 <- pivot_wider(data, id_cols=c(lizard,temp,cort,species,food_ingested),
 names_from = Day, values_from = mass)

# Estimate mass increase and rescale it to avoid negative values
data_2 <- data_2 %>% 
 mutate(delta_mass=After-Before) %>%
 mutate(rescaled_delta_mass = ifelse(is.na(delta_mass),
  NA, 
  delta_mass + abs(min(delta_mass, na.rm = T)) +1 ))
write.csv(data_2, file = "./data/Stressor_2.csv")

# Split new database by species
deli_stressor_2 <- data_2 %>% filter (species == "delicata")
guich_stressor_2 <- data_2 %>% filter (species == "guichenoti")

# Explore distribution of mass increase
  # A) L.deli
plot_mass <- ggplot(deli_stressor_2, aes(x = rescaled_delta_mass, fill = interaction(as.factor(temp),as.factor(cort), sep="_"))) +
  geom_histogram(position = "identity", alpha = 0.5, bins = 30, color = "black") +
  scale_fill_manual(values = c("28_A" = "red", "23_A" = "blue", "28_B" = "salmon", "23_B" = "lightblue")) +
  labs(x = "Mass", y = "Density", fill = "temp ~ cort") +
  theme_minimal() +
  facet_grid(as.factor(deli_stressor_2$cort)~as.factor(deli_stressor_2$temp), scales = "free_y")
print(plot_mass)

mass_normal_test<-by(deli_stressor_2$rescaled_delta_mass, interaction (deli_stressor_2$cort, deli_stressor_2$temp), function(x) shapiro.test((x)))
print(mass_normal_test)

  #B) L.guich
plot_mass <- ggplot(guich_stressor_2, aes(x = rescaled_delta_mass, fill = interaction(as.factor(guich_stressor_2$temp),as.factor(guich_stressor_2$cort), sep="_"))) +
  geom_histogram(position = "identity", alpha = 0.5, bins = 30, color = "black") +
  scale_fill_manual(values = c("28_A" = "red", "23_A" = "blue", "28_B" = "salmon", "23_B" = "lightblue")) +
  labs(x = "Mass", y = "Density", fill = "Temperature ~ CORT") +
  theme_minimal() +
  facet_grid(guich_stressor_2$cort~guich_stressor_2$temp, scales = "free_y")
print(plot_mass)

mass_normal_test<-by(guich_stressor_2$rescaled_delta_mass, interaction (guich_stressor_2$cort, guich_stressor_2$temp), function(x) shapiro.test((x)))
print(mass_normal_test)

@@

```{r, models}
#| label: models
#| echo: false
#| warning: false

# Delta mass L. delicata
  deli_mod <- brm(rescaled_delta_mass ~ food_ingested+cort*temp, data = deli_stressor_2, family = gaussian(link = "identity"), chains = 4, cores = 4, iter = 2000, warmup = 1000, control = list(adapt_delta = 0.99))
  plot(deli_mod)
  summary(deli_mod)
# Delta mass L. guichenoti
  guich_mod <- brm(rescaled_delta_mass ~ food_ingested+cort*temp, data = guich_stressor_2, family = gaussian(link = "identity"), chains = 4, cores = 4, iter = 2000, warmup = 1000, control = list(adapt_delta = 0.99))
  summary(guich_mod)
bayes_R2(guich_mod)

      # Seems that only food ingestion is significant in both species

# Frozen behaviour L. delicata
  deli_mod <- brm(log(lat_to_move+1) ~ Day*cort*temp + (1|lizard), data = deli_stressor, family = gaussian(link = "identity"), chains = 4, cores = 4, iter = 2000, warmup = 1000, control = list(adapt_delta = 0.99))
  summary(deli_mod)
  plot(deli_mod)
bayes_R2(deli_mod)

# Frozen behaviour L. delicata
 guich_mod <- brm(log(lat_to_move+1) ~ Day*cort*temp + (1|lizard), data = guich_stressor, family = gaussian(link = "identity"), chains = 4, cores = 4, iter = 2000, warmup = 1000, control = list(adapt_delta = 0.99))
  summary(guich_mod)
  plot(guich_mod)
bayes_R2(guich_mod)

@@

```{r, plotsMass}
#| label: plotsMass
#| echo: false
#| warning: false

# Plot delta Mass L. delicata
plot_deli <- ggplot(deli_stressor_2, aes(x = food_ingested, y = rescaled_delta_mass, color=interaction(cort,temp))) +  
  stat_smooth(method = "glm", formula = y ~ x, se = TRUE, linewidth=2, alpha = 0.1) +
  scale_color_manual(values = c("CORT.Cold"="black", "Control.Cold"="grey", "CORT.Hot"="orange2", "Control.Hot"="moccasin"),
    labels=c("CORT-Cold (n=12)", "Control-Cold (n=12)", "CORT-Hot (n=11)", "Control-Hot (n=12)")
  ) +
  labs(y = "Increase in mass (mg)", x = "Ingestion rate (n items)", color="Treatment") +
  theme_classic() +
  theme(
    axis.title = element_text(size = 18, family = "sans"),  
    axis.text = element_text(size = 10, family = "sans"),  
    legend.title = element_text(size = 14, family = "sans"), 
    legend.text = element_text(size = 12, family = "sans")  
  )
print(plot_deli)
ggsave("./output/figures/fig_deli_Mass.png", plot=plot_deli, width = 18, height = 9, units = "cm", dpi = 3000)
# Plot delta Mass L. guichenoti
plot_guich <- ggplot(guich_stressor_2, aes(x = food_ingested, y = rescaled_delta_mass, color=interaction(cort,temp))) +  
  stat_smooth(method = "glm", formula = y ~ x, se = TRUE, linewidth=2, alpha = 0.1) +
  scale_color_manual(values = c("CORT.Cold"="black", "Control.Cold"="grey", "CORT.Hot"="orange2", "Control.Hot"="moccasin"),
    labels=c("CORT-Cold (n=10)", "Control-Cold (n=7)", "CORT-Hot (n=11)", "Control-Hot (n=10)")
  ) +
  labs(y = "Increase in mass (mg)", x = "Ingestion rate (n items)", color="Treatment") +
  theme_classic() +
  theme(
    axis.title = element_text(size = 18, family = "sans"),  
    axis.text = element_text(size = 10, family = "sans"),  
    legend.title = element_text(size = 14, family = "sans"), 
    legend.text = element_text(size = 12, family = "sans")  
  )
print(plot_guich)
ggsave("./output/figures/fig_guich_Mass.png", plot=plot_guich, width = 18, height = 9, units = "cm", dpi = 3000)
```

 ```{r, plotsFrozenbehaviour}
#| label: plotsFrozenbehaviour
#| echo: false
#| warning: false

# Plot time frozen L. delicata
plot_deli_BEH <- ggplot(deli_stressor, aes(x = Day, y = lat_to_move, fill=interaction(cort,temp))) +  
  geom_boxplot(outlier.shape = NA) +
  ylim(0, 450) +
  scale_fill_manual(values = c("CORT.Cold"="black", "Control.Cold"="grey", "CORT.Hot"="orange2", "Control.Hot"="moccasin"),
    labels=c("CORT-Cold (n=12)", "Control-Cold (n=12)", "CORT-Hot (n=11)", "Control-Hot (n=12)")
  ) +
  labs(y = "Time frozen (s)", x = "Day", fill="Treatment") +
  theme_classic() +
  theme(
    axis.title = element_text(size = 18, family = "sans"),  
    axis.text = element_text(size = 12, family = "sans"),  
    legend.title = element_text(size = 14, family = "sans"), 
    legend.text = element_text(size = 12, family = "sans")  
  )
print(plot_deli_BEH)
ggsave("./output/figures/fig_deli_BEH.png", plot=plot_deli_BEH, width = 15, height = 12, units = "cm", dpi = 3000)

# Plot time frozen L. guichenoti
plot_guich_BEH <- ggplot(guich_stressor, aes(x = Day, y = lat_to_move, fill=interaction(cort,temp))) +  
  geom_boxplot(outlier.shape = NA) +
  ylim(0, 450) +
  scale_fill_manual(values = c("CORT.Cold"="black", "Control.Cold"="grey", "CORT.Hot"="orange2", "Control.Hot"="moccasin"),
    labels=c("CORT-Cold (n=10)", "Control-Cold (n=7)", "CORT-Hot (n=11)", "Control-Hot (n=10)")
  ) +
  labs(y = "Time frozen (s)", x = "Day", fill="Treatment") +
  theme_classic() +
  theme(
    axis.title = element_text(size = 18, family = "sans"),  
    axis.text = element_text(size = 12, family = "sans"),  
    legend.title = element_text(size = 14, family = "sans"), 
    legend.text = element_text(size = 12, family = "sans")  
  )
print(plot_guich_BEH)
ggsave("./output/figures/fig_guich_BEH.png", plot=plot_guich_BEH, width = 15, height = 12, units = "cm", dpi = 3000)
